### 2014-10-08 Writeup/paper_v3.tex
significant rewrite of conclusions

\documentclass[twocolumn,useAMS,usenatbib]{mn2e}
\usepackage{natbib}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{epsfig,graphicx}
\usepackage{subfig}
\usepackage{float}
\usepackage{color}
\usepackage{hyperref}
\graphicspath{{Figures/}}
\topmargin-1cm

% For correct printing on US Letter, while still working on A4
\topmargin-1cm

\newcommand{\rachel}[1]{{\textcolor{red}{#1}}}
%\newcommand{\rachel}[1]{}
\newcommand{\arun}[1]{{\textcolor{blue}{#1}}}
%\newcommand{\arun}[1]{}
\newcommand{\claire}[1]{{\textcolor{green}{#1}}}
\def\a{\alpha}
\def\reff@jnl#1{{\rm#1\/}}

\def\aj{\reff@jnl{AJ}}                  % Astronomical Journal
\def\araa{\reff@jnl{ARA\&A}}            % Annual Review of Astron and Astrophys
\def\apj{\reff@jnl{ApJ}}                        % Astrophysical Journal
\def\apjl{\reff@jnl{ApJ}}               % Astrophysical Journal, Letters
\def\apjs{\reff@jnl{ApJS}}              % Astrophysical Journal, Supplement
\def\apss{\reff@jnl{Ap\&SS}}            % Astrophysics and Space Science
\def\aap{\reff@jnl{A\&A}}               % Astronomy and Astrophysics
\def\aapr{\reff@jnl{A\&A~Rev.}}         % Astronomy and Astrophysics Reviews
\def\aaps{\reff@jnl{A\&AS}}             % Astronomy and Astrophysics, Supplement
\def\baas{\reff@jnl{BAAS}}              % Bulletin of the AAS
\def\jrasc{\reff@jnl{JRASC}}            % Journal of the RAS of Canada
\def\memras{\reff@jnl{MmRAS}}           % Memoirs of the RAS
\def\mnras{\reff@jnl{MNRAS}}            % Monthly Notices of the RAS
\def\physrep{\reff@jnl{Phys.Rep.}}
\def\pra{\reff@jnl{Phys.Rev.A}}         % Physical Review A: General Physics
\def\prb{\reff@jnl{Phys.Rev.B}}         % Physical Review B: Solid State
\def\prc{\reff@jnl{Phys.Rev.C}}         % Physical Review C
\def\prd{\reff@jnl{Phys.Rev.D}}         % Physical Review D
\def\prl{\reff@jnl{Phys.Rev.Lett}}      % Physical Review Letters
\def\pasp{\reff@jnl{PASP}}              % Publications of the ASP
\def\pasj{\reff@jnl{PASJ}}              % Publications of the ASJ
\def\skytel{\reff@jnl{S\&T}}            % Sky and Telescope
\def\solphys{\reff@jnl{Solar~Phys.}}    % Solar Physics
\def\sovast{\reff@jnl{Soviet~Ast.}}     % Soviet Astronomy
\def\ssr{\reff@jnl{Space~Sci.Rev.}}     % Space Science Reviews
\def\nat{\reff@jnl{Nature}}             % Nature

\def\sun{\hbox{$\odot$}}
\def\farcs{\hbox{$.\!\!^{\prime\prime}$}}
\newcommand{\hvol}{h^{3}{\mathrm{Mpc}}^{-3}}
\newcommand{\hmpc}{\ensuremath{h^{-1}\mathrm{Mpc}}}
\newcommand{\hkpc}{\ensuremath{h^{-1}\mathrm{kpc}}}
\newcommand{\hMsun}{h^{-1}M_{\odot}}
\newcommand{\Msun}{M_{\odot}}
\newcommand{\kms}{{\,{\rm km}\,{\rm s}^{-1}}}
\newcommand{\Omegam}{\Omega_{m}}
\newcommand{\Omegab}{\Omega_{b}}
\newcommand{\Omegal}{\Omega_{\Lambda}}
\newcommand{\xig}{\xi_{\rm gg}(r)}
\newcommand{\xih}{\xi_{\rm hh}(r)}
\newcommand{\xim}{\xi_{\rm mm}(r)}
\newcommand{\ds}{\ensuremath{\Delta\Sigma}}
\newcommand{\scinv}{\ensuremath{\Sigma_c^{-1}}}
\newcommand{\avgscinv}{\ensuremath{\langle\Sigma_c^{-1}\rangle}}
\newcommand{\hinvk}{$h^{-1}$kpc}
\newcommand{\avgnm}{\ensuremath{\langle N(M)\rangle}}
\newcommand{\fclust}{\ensuremath{f_\mathrm{clust}}}
\newcommand{\fbcg}{\ensuremath{f_\mathrm{BCG}}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}

% upright d in integrals
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\putcite}{\textbf{(CITE)}}
\newcommand{\ic}{\ensuremath{I_\mathrm{C}}}
\newcommand{\pc}{\ensuremath{G_\mathrm{C}}}
\newcommand{\ps}{\ensuremath{G_\mathrm{T}}}
\newcommand{\tic}{\ensuremath{\tilde{I}_\mathrm{C}}}
\newcommand{\tpc}{\ensuremath{\tilde{G}_\mathrm{C}}}
\newcommand{\tps}{\ensuremath{\tilde{G}_\mathrm{T}}}
\newcommand{\tkpd}{\ensuremath{\tilde{T}}}
\newcommand{\ticpd}{\ensuremath{\tilde{P}}}
\newcommand{\ticpdr}{\ensuremath{\tilde{P}^\mathrm{(rot)}}}
\newcommand{\ticpds}{\ensuremath{\tilde{P}^{(\gamma)}}}
\newcommand{\ticpdrs}{\ensuremath{\tilde{P}^{(\mathrm{rot},\gamma)}}}
\newcommand{\tks}{\ensuremath{\tilde{K}_\mathrm{T}}}
\newcommand{\tksr}{\ensuremath{\tilde{K}_\mathrm{T}^\mathrm{(rot)}}}
\newcommand{\tis}{\ensuremath{\tilde{I}^{(\gamma)}_\mathrm{T}}}
\newcommand{\tisr}{\ensuremath{\tilde{I}^{(\mathrm{rot},\gamma)}_\mathrm{T}}}
\newcommand{\is}{\ensuremath{I_\mathrm{T}}}
\newcommand{\isr}{\ensuremath{I_\mathrm{T}^\mathrm{(rot)}}}
\newcommand{\rpix}{\ensuremath{R_\mathrm{pix}}}
\newcommand{\nimg}{\ensuremath{N_\mathrm{img}}}
\newcommand{\nrand}{\ensuremath{N_\mathrm{rand}}}
\newcommand{\nset}{\ensuremath{N_\mathrm{set}}}
\newcommand{\nc}{\ensuremath{N_\mathrm{C}}}
\newcommand{\ns}{\ensuremath{N_\mathrm{S}}}
\newcommand{\nt}{\ensuremath{N_\mathrm{T}}}
\newcommand{\erms}{\ensuremath{e_\mathrm{rms}}}
\newcommand{\etot}{\ensuremath{e_\mathrm{tot}}}
%\newcommand{\newtext}{\emph}
\newcommand{\newtext}{}
\newcommand{\reftext}[1]{\textbf{#1}}
%\newcommand{\arcsec}{\ensuremath{^{\prime\prime}}}

%More new commands by Arun Kannawadi
\newcommand{\mg}{\ensuremath{M_G}}
\newcommand{\mi}{\ensuremath{M_I}}
\newcommand{\sersicn}{S\'{e}rsic $n$ }
\newcommand{\sersic}{S\'{e}rsic }
\newcommand{\btt}{Bulge-to-Total }
\newcommand{\s}{\ensuremath{\mathcal{S}}}
\newcommand{\scinot}[2]{\ensuremath{#1 \times 10^{#2}}}

\title[WL simulation]{The impact of cosmic variance on simulating weak lensing surveys}

\author[Kannawadi et al.]
{Arun Kannawadi$^1$\thanks{\tt akannawa@andrew.cmu.edu}, 
Rachel Mandelbaum$^1$,
Claire Lackner$^2$, 
\\$^1$McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15217, USA
\\$^2$Kavli Institute for the Physics and Mathematics of the Universe (WPI), Todai Institutes for Advanced Study,\\ the University of Tokyo, Kashiwa, Japan.
}

\date{\today}

\begin{document}
\bibliographystyle{mn2e2}
\maketitle

\begin{abstract}
According to our current understanding, galaxy shapes and morphologies should depend on various factors such as the local environment. Realistic image simulations for calibration of weak lensing analysis methods that use training samples from the Hubble Space Telescope can therefore be affected by these trends, due to the limited volume of the universe that has been surveyed by Hubble. We will show how redshift slices in a volume-limited subsample of COSMOS can be classified as overdense or underdense (or neither), and how the statistical properties of various morphological parameters such as ellipticity, \sersicn, \btt ratio and color differ in these bins. This study requires a careful distinction between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. We conclude with some guidance for how upcoming surveys can use COSMOS data as the basis for weak lensing simulations without having their conclusions overly affected by cosmic variance.  
\rachel{I have some comments on this abtract, but prefer to do all
  revision of the abstract at the end once the paper is finalized.}
\end{abstract}

\begin{keywords}
 Gravitational lensing: weak --- Cosmology: Large-scale structure of Universe --- Galaxies: evolution.
\end{keywords}

\section{Introduction}
\label{S:intro}

Weak gravitational lensing, the deflection of light by mass, is one of the cleanest ways to study the nature of dark energy by tracking the growth of structure in the Universe as a function of time \citep[e.g.,][]{2001PhR...340..291B,2006astro.ph..9591A,2013PhR...530...87W}.
As light from background sources passes by matter (including dark matter) on its way to us, the apparent shapes of the background galaxies get distorted, and the galaxies get slightly magnified as well.
Because of its sensitivity to dark matter and dark energy, major surveys such as the Hyper Suprime-Cam \citep[HSC;][]{HSC_Overview}, Dark Energy Survey~\citep[DES;][]{DESC}, the KIlo-Degree Survey~\citep[KIDS;][]{KIDS}, the Panoramic Survey Telescope and Rapid Response System~\citep[PanSTARRS;][]{PanSTARRS_2010},
the Large Synoptic Survey Telescope~\citep[LSST;][]{LSST_Book}, Euclid\footnote{\url{http://sci.esa.int/euclid/}, \url{http://www.euclid-ec.org}} \citep{EuclidReport},
%(\rachel{cite http://adsabs.harvard.edu/abs/2011arXiv1110.3193L}),
and Wide-Field Infrared Survey Telescope~\citep[WFIRST;][]{WFIRST_Final}
%(\rachel{for WFIRST, you should actually cite the Spergel et al report that was on the arxiv in the past year, not Green et al})
are planned for the next two decades to gather enormous quantities of weak lensing data that will lead to precise constraints on the growth of structure with time, and therefore cosmological parameters.
%Weak lensing effects need to be removed from the Cosmic Microwave Background (CMB) analysis to understand the initial conditions of the universe.

For the upcoming surveys to achieve their promise, their systematic error budgets must be below their statistical error budgets.
Systematic error budgets for weak lensing surveys typically include
astrophysical effects, such as intrinsic alignments of galaxy shapes
with large scale density fields \citep[e.g.,][]{2014arXiv1407.6990T} and the effect of baryons on the
matter power spectrum \citep[e.g.,][]{2011MNRAS.415.3649V}, as well as observational uncertainties such as
the ability to robustly infer shears from galaxy observed shapes or
photometric redshifts from their observed colors. 
Given the expected sub-per cent errors on upcoming surveys, systematic errors must be reduced from
 their typical level in the current state-of-the-art
 measurements that typically achieve $\sim 5$ per cent statistical
 errors at best
 \citep[e.g.,][]{2010A&A...516A..63S,2013MNRAS.432.2433H,2013ApJ...765...74J,2013MNRAS.432.1544M}.

One method that is commonly used to test for the presence of
systematic errors in the shear estimation process is image simulation,
where we can cleanly test whether our methods of shear estimation
recover the ground truth. This is a valuable test, considering the
numerous sources of additive and multiplicative bias such as a
mismatch between galaxy model assumptions and actual galaxy light
profiles \citep[e.g.,][]{2010MNRAS.404..458V, 2010A&A...510A..75M}, biases due to the effects of pixel noise on the
shear estimates
\citep{2012MNRAS.427.2711K,2012MNRAS.424.2757M,2012MNRAS.425.1951R},
and ellipticity gradients \citep{2010MNRAS.406.2793B}.  These biases
often differ for galaxies with different morphologies (e.g., disks
vs. ellipticals), sizes, $S/N$, and shape \citep{2010MNRAS.405.2044B,2012MNRAS.423.3163K}. 
A general requirement for simulations used to test shear recovery is
that they should be as realistic as possible.

 Realistic simulations may use samples based on images from the Hubble
 Space Telescope ({\em HST}).
Software packages like
{\sc GalSim}\footnote{\url{https://github.com/GalSim-developers/GalSim}}
\citep{2014arXiv1407.7676R} can generate images of galaxies from the
{\em HST} as they would appear with an additional lensing shear and
viewed by some lower resolution telescope.  Examples of training samples from the {\em HST} include 
the COSMOS survey \citep[used by the GREAT3 challenge,][]{great3} or the
 Ultra Deep Field \citep[UDF, used by][]{2013ApJ...765...74J}.  These
 two examples serve as the extremes in the {\em HST} samples used as
 the basis for image simulation, with COSMOS being shallower but
representing the current widest contiguous area surveyed by the {\em HST}, and the
 latter being extremely deep but narrow.

For a variety of physical reasons, some of which are still not fully
understood, the shape and morphology of
galaxies depends on their local environment
\citep[e.g.,][]{2014arXiv1402.1172C,2014MNRAS.444.2200D}. 
Hence, local overdensities or underdensities
observed in these {\em HST} fields may (given the small size of the field)
cause the properties of the galaxy population in redshift slices to be
atypical depending on the environment in that slice.  This would have
the undesired consequence of including variation in galaxy properties
due to the COSMOS (or other) survey cosmic variance in the simulated galaxy sample in that redshift slice, rather than
only including ensemble effects that would appear in a large cosmological volume, such as true redshift evolution of
galaxy properties.  Our goal is to quantify the degree to which the
morphology-density correlations in COSMOS cause noticeable changes in
the galaxy populations in narrow redshift slices at a level that could
result in difficulty using the sample to derive redshift-dependent
shear calibrations.   Upcoming surveys will study lensing as a function of
redshift and therefore need to simulate galaxy samples at different
redshifts in order to assess the shear calibration at each redshift.

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
The paper is structured as follows: in Sec.~\ref{S:data}, we describe
the data that we use for this study.  In Sec.~\ref{S:methods}, we
describe our methods for deriving the relevant galaxy properties
like environment, morphology, and shape.  
Using these ingredients, we present our results in Sec.~\ref{S:results}
and discuss their implications in Sec.~\ref{S:summary}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
The COSMOS survey \citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).  We use the COSMOS survey to
define a parent sample of galaxy images to be used for making image
simulations, following the approach taken to this problem in
\cite{2012MNRAS.420.1518M,great3}.

We apply the following set of initial cuts to the COSMOS data, the
first two of which
are motivated and explained in more detail by \cite{COSMOS_Alexie}:
\begin{enumerate}
 \item \texttt{MU\_CLASS=1}: This criterion uses a comparison between
   the peak surface brightness and the background level to achieve a
   robust star/galaxy separation, with galaxies having \texttt{MU\_CLASS=1}. 
 \item \texttt{CLEAN=1}: Objects near bright stars or those containing
   saturated pixels were removed; the rest pass this cut on \texttt{CLEAN}. 
% \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
%                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
%                                       \arun{copied from Mandelbaum
%                                         2011.}
 \item \texttt{GOOD\_ZPHOT\_SOURCE =1}: This cut requires that
   photometric redshifts be reliable and good enough to draw
   conclusions about the population (see \citealt{2012MNRAS.420.1518M}
   for details).
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

%Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
%is $0.12''$. 
High resolution images taken through the wide F814W filter (broad
\emph{I}) for all galaxies passing the above cuts were used to create
a collection of postage stamp images for the GREAT3 challenge
\citep{great3}, using the procedure described in
\cite{2012MNRAS.420.1518M}.  Each galaxy postage stamp image has a
corresponding PSF image that can be used by {\sc GalSim} or other
software to remove the effects of the {\em HST} PSF before simulating
the galaxy image as it would appear at lower resolution.

To better characterize the galaxy population, parametric models were
fit to the light profiles of these galaxies.  These were carried out
using the method described in \cite{Claire_Fits}, and include
\sersic profile fits and 
2 component bulge $+$ disk fits described in detail in \cite{great3}
and briefly in Sec.~\ref{sub:axisratio} of this work.

In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam, the
Canada-French Hawaii Telescope (CFHT) and KPNO/CTIO, yielding many bands
of imaging data from which to determine high-fidelity photometric
redshifts. 
Photometric redshifts were determined by
\cite{COSMOS_Photoz_30band}. The accuracy of photometric redshifts for
$m_\text{F814W}\le$ 22.5 is $\sigma_{\Delta z} = 0.007(1+z)$; for $m_\text{F814W}\le$ 24, $\sigma_{\Delta z} = 0.012(1+z)$.
The photometric redshift values become noisier beyond $z\sim 1.2$, and
the  fits to the galaxy light profiles are also somewhat noisy once we
go beyond $m_\text{F814W}\sim 23.5$.   
For this reason, we will exclude all galaxies that have F814W
magnitude fainter than 23.5. However, we will use the
$m_\text{F814W}\le 25.2$ sample that was generated for the GREAT3
challenge to estimate the completeness, which is useful when
generating a volume-limited sample (Sec.~\ref{sub:volumelimiting}).
We first use the $z<1$ flux-limited sample to fit parametric redshift
distribution models (Sec.~\ref{sub:overdensities}), and then
restrict ourselves to $z\le1$ sample for all further analysis. 
% ($S/N \sim 50$ {\bf ?} )
% \rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
%   not the older Mobasher ones.  Need to fix citations. Also, we care
%   about the photo-$z$ accuracy beyond 22.5, so it's important to give
%   some information about it to the depth that we actually use.  Finally, I
%   recommend against mentioning the $S/N$ because the calculation is
%   complicated by the correlated noise, so most people use use
%   magnitude instead of $S/N$.}

Stellar mass estimates were obtained~\citep{COSMOS_XRAY} using the
Bayesian code described in~\cite{KEVIN_MSTAR}.  This process involves
constructing a grid of models that vary in age, star formation
history, dust content and metallicity (always assuming a Chabrier IMF;
\citealt{ChabrierIMF}), 
 to which the observed galaxy spectral energy distributions (SEDs) and photometric redshift are compared. At each grid point, the probability that the 
SED fits the model is calculated, and by marginalizing over the nuisance parameters in the grid, the stellar mass probability distribution is obtained. The median of this distribution
is taken as the stellar mass estimate.

%We will analyze how the galaxy morphology and intrinsic shape depends on
%the environment in which they reside.

\section{Methods}
\label{S:methods}

In order to study the variation in the intrinsic ellipticity
distribution and various morphological indicators with the galaxy environment, there are three main steps to be carried out:
\begin{enumerate}
 \item Identify overdense and underdense environments in our survey
   from the redshift distribution of galaxies (Sec.~\ref{sub:overdensities});
 \item volume-limit the sample such that Malmquist bias is minimized
   before comparing galaxies in different redshift slices
   (Sec.~\ref{sub:volumelimiting}); and 
 \item estimate the galaxy axis ratios and other morphological indicators such as  \sersic index and bulge-to-total ratios (Sec.~\ref{sub:axisratio}).
\end{enumerate}

In this section we will describe how these steps were carried out.

%\rachel{Here you should put a few introductory sentences about the purpose of this section. Could be something like ``To study the  variation in the intrinsic ellipticity distribution with environment, there are two main steps.  The first is to identify redshift ranges with overdense and underdense environments.  The second is to estimate galaxy axis ratios.  In this section we describe how these two steps were done.''}

\subsection{Finding overdensities}
\label{sub:overdensities}
\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_fluxlimited}
  %\caption{\rachel{You show this figure twice.  Why?  The legend needs to say the name of the distribution, not show the functional form.}}
  \caption{Upper panel: Redshift distribution of flux-limited
    ($m_\text{F814W}\le 23.5$) sample with photometric redshift bins
    that are 0.05 wide. Two analytical functions, $p_1(z)$ and $p_2(z)$, defined in Eqns.~\ref{E:pz1} and~\ref{E:pz2} with best fit
    parameters are plotted over it. 
           Lower panel: Plot of $\delta_{g,\text{1D}} =
           N/N_{\text{mod}}-1$ with each functional form as the model
           for each redshift bin. \arun{Note to self: Give parameter values.}
\rachel{How does this figure look if you show the plot range with the
  histogram going to $z=0$ (still only fitting to $0.3<z<1.25$)?  It
  might be nicer to do that and put a vertical line at $z=0.3$ to
  remind the reader that you only fit above that redshift.}
}
  \label{fig:redshift_fluxlimited}
\end{figure}

%\rachel{I am a bit confused here.  Our initial strategy was as follows: use all galaxies with $F814W<23.5$ or 25.2 to define a redshift distribution that we fit with one of our functional forms, use that to define overdensities, and then later when we're checking
%  the variation of properties with redshift, we restrict to
%  volume-limited subsamples.  But here you seem to be volume-limiting
%  as the first step, which is not really appropriate. The functional
%  forms that people have derived for surveys are predicated on the
%  assumption that they are flux-limited.  What is the justification
%  for volume-limiting first?  if there is none, then we should not be
%  doing it. And then this section will have 3 subsection: finding
%  overdensities, estimating axis ratios, and volume limiting,
%  where we need to explain the two choices (using $B$-band evolution
%  and using no evolution).}

It is important to keep in mind when considering the environment
estimation that our goal is not to create a full 3D mapping of the
density field within the COSMOS region (a task that was already
addressed by \citealt{Kovac_Density10k} using the zCOSMOS
spectroscopic sample).  Instead, we make a coarse, 1D division of the
COSMOS survey into redshift slices, just as would be done when making
galaxy redshift slices as input to a weak lensing survey simulation.
For each redshift slice, we can then check whether the environment is
overdense or underdense on average.  Our approach will tend to wash
out some real trends from a 3D study, but is appropriate given our
scientific goal of testing effects of the environment on weak lensing
simulations based on the COSMOS survey.

%In a wide-field survey, regions of overdensities are selected  identified by computing 3-dimensional comoving densities. But in narrow field surveys like COSMOS, previous work {\bf cite } have shown
%that regions of overdensities can be identified from a 1-dimensional histogram of redshifts, at least for low redshifts. 
%\rachel{This is not entirely the point.  As we've discussed, the point
%  is that we do not care about 3d densities for this particular science.  Weak lensing simulations
%  will use galaxies selected in 2d slices, and hence we will simply do
 % it that way as well.  But we use the comparison with a 3d analysis
 % as a test of our method.}

%\rachel{I thought we were only going up to $z=0.85$ but here you say $1$?}
%For our sample of galaxies, volume-limited upto $z=1$ by imposing an $M_I$ cut, we fit parametric models to the histogram of photometric redshifts in order to assign values of overdensity. The figures show chi-squared and gamma functions fitted to the histogram \citep{Redshift_modelling}.
For our (flux-limited) sample of galaxies, up to $z=1.0~(1.25)$
(\rachel{fix this to the correct one once we make final decisions}), we fit
parametric models to the histograms of photometric redshifts in order
to assign values of overdensity.  We choose our bins to be $0.05$ wide
starting from $z=0.3$, where the bin width is selected to be somewhat
larger than the photometric redshift error but narrow enough that we
can still identify rather than averaging over real cosmological
structures.  We neglect the lowest
redshifts which have negligible cosmological volume and where the
galaxy population tends to be intrinsically bright and large enough
that a non-negligible fraction is lost due to the cuts we impose
(Sec.~\ref{S:data}).

The parametric redshift distributions that we use are
\begin{equation}\label{E:pz1}
 p_1(z) \propto z^{a-1}\exp\left[{-bz}\right]
\end{equation}
and 
\begin{equation}\label{E:pz2}
 p_2(z) \propto z^2\exp{\left[-\left(\frac{z}{z_0}\right)^{1.5}\right]},
\end{equation}
the latter of which was first presented by
\cite{Redshift_modelling}. Here $a$, $b$, and
$z_0$ are free parameters that are to be determined.  The
normalization constants depend not only on the parameters but also on
the lower and the upper limit of the redshifts considered, where we
fix the normalization to ensure that the predicted number of galaxies
in the range used ($0.3<z<1$) is equal to the actual number. 
Fig.~\ref{fig:redshift_fluxlimited} shows the photometric redshift
histogram together with the best-fitting parametric distributions.
\rachel{New suggestion re: 2nd redshift distribution: it just looks so
  bad, I worry that this functional form is simply wrong 
  and we should use something newer that has been shown to work with
  modern redshift surveys.  Let me suggest something you can
  do as an immediate sanity check.   Look at
  \cite{2004ApJ...617..765C} table 4, which gives parametrized fits to
  the redshift distribution from DEEP2 for samples binned by
  $R_\text{AB}$ and $I_\text{AB}$.  Use the results for the
  $18<I_\text{AB}<23$ bin and the $18<I_\text{AB}<24$ bin,
  interpolating linearly between them since your sample is limited at
  $23.5$.  Their results suggest that either $z^2 \exp{[-z/0.262]}$ or
  $z^2 \exp{[-(z/0.361)^{1.2}]}$ should be decent descriptions of your sample.  Just as a quick
  sanity check can you please plot those  distributions (appropriately
  normalized) on top of what you have in the top panel of figure 1,
  and see how they look?  If they look OK, then I would be inclined to
  completely eliminate the 2nd distribution, and just use the 1st one
  that you're fitting and one of the ones from Coil as your way of
  checking overdensities and underdensities.
}

The estimated overdensity in a redshift bin is defined by comparing
the observed galaxy counts in the bin with the counts that are
predicted in that bin by one of the models in Eqs.~\eqref{E:pz1}
and~\eqref{E:pz2}:
\begin{equation}
\delta_{g,\text{1D}}=\frac{(N-N_{\text{mod}})}{N_{\text{mod}}},
\end{equation}
where
\begin{equation}
N_\text{mod} = \int_{z_\text{min}}^{z_\text{max}} p(z) \,\mathrm{d}z
\end{equation}
is determined by integrating the redshift distribution within the
limits of that redshift slice. 
%(\rachel{This $\delta$ definition is important, so it should be in an equation environment.})
Note that $\delta_{g,\text{1D}}$ is dependent on our choice of model
redshift distribution, and should have a mean value of $0$ over the
entire redshift range. % If a model and the distribution agree, then $\delta_{g,\text{1D}}=0$ for that model at all the redshifts.

Our decision criterion for identifying overdense and underdense
redshift slices involves leaving a 10 per cent margin around an
overdensity of zero; i.e., if $\left| \delta_{g,\text{1D}} \right| <
0.1$, that is considered ``neutral'' (neither overdense nor
underdense on average).  We can then label each redshift slice as
either overdense, underdense, or neutral as follows: 
We label a redshift bin as overdense if at least one model gives a
value of $\delta_{g,\text{1D}} > 0.1$ while the other gives
$\delta_{g,\text{1D}} > -0.1$ 
(neutral or overdense), and vice versa for the underdense regions. We
label a redshift bin as neutral if both models give
$\delta_{g,\text{1D}}$ within the neutral region,  
\emph{or} if use of one model redshift distribution results in the
conclusion that the bin is overdense while the over leads to the
conclusion that it is underdense.
We thus identify the regions $z=0.30-0.40$, $0.65-0.75$, and
$0.80-0.85$ as overdense; $z=0.55-0.65$ and $0.75-0.80$ as underdense; and $z=0.40-0.55$ as unclassified. 
%Although, the $z=0.40-0.55$ bin appears to be underdense, for reasons mentioned in \S\ref{sub:volumelimiting}, we will not call them underdense.

We have adopted this purely 1D environment classification for reasons
explained at the beginning of this section. However, as a sanity check
we can compare it with a more rigorous study that includes information
about structure in the plane of the sky.  
 \cite{Kovac_Density10k} used a sample of $\sim$10~000 zCOSMOS
 spectroscopic galaxies with $I_{AB}<22.5$ to reconstruct the three dimensional overdensity field up to $z\sim 1$.
We find that our classification of overdensities and underdensities
agrees with this work, except for our two highest redshift bins.
We believe that this disagreement is due to the errors in our
photometric redshifts, with the overdensity reported by
\cite{Kovac_Density10k} in the $z=0.875-1$ range 
leaking into our $z=0.80-0.85$ slice.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{hist2d_mag_mi}
   \caption{2D histogram of galaxies in apparent magnitude
     ($m_\text{F814W}$) and absolute magnitude ($M_I$) in the redshift range $[0.3, 1.0]$. The blue dotted line with a unit slope is a boundary at $z=1$, fitted by eye.
     The green vertical dotted line represents our luminosity cut. } 
   \label{fig:2Dhist}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{MAG_histograms}
   \caption{Distribution of absolute magnitude $M_I$ for various
     flux-limited samples in the redshift range $[0.80, 0.85]$ are plotted together. The vertical line
     corresponds to our luminosity cut of $-20.8$, brighter than which
     the $m_\text{F814W}<23.5$ sample includes $>95.3$ per cent of the
     galaxies in the $m_\text{F814W}<25.2$ sample.
   }
   \label{fig:MIhist}
 \end{figure}

\subsection{Volume limiting}
\label{sub:volumelimiting}

COSMOS is a flux-limited survey and is therefore affected by Malmquist
bias, with the galaxy samples at higher redshift being intrinsically
brighter on average.  
Our analysis involves comparing galaxies in different redshift slices
to identity significant differences in morphology that arise due to
morphology-density correlations.  Such an analysis would be very
difficult with a flux-limited sample because there would be some
variation in morphology with redshift just due to the intrinsic change
in the sample properties.  For a fair comparison, we must restrict
ourselves to galaxies that are bright enough that they would be
observed at all redshifts that we consider, which is achieved by
volume-limiting the sample.  We consider three different ways of
carrying out this process, which results in three different galaxy
sample selections, all of which we will use in the remainder of the
analysis.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}

Our first approach is to generate a volume-limited sample that is
complete up to $z=1$, by applying a cut on luminosity such that only
galaxies instrinsically brighter than a certain threshold (determined
in detail below) is considered. This threshold is set on the
$k$-corrected $I$-band absolute magnitudes ($M_I$) from the COSMOS PSF-matched
photometry catalog.   The joint distribution of $m_\text{F814W}$ and
$M_I$ is shown in Fig.~\ref{fig:2Dhist}. 
Since the parent sample contains fainter galaxies and is quite
complete to $m_\text{F814W}=25.2$, we compare the $M_I$ distribution
of the $m_\text{F814W}=23.5$ sample with flux-limited samples that
have fainter flux limits, to see where the $m_\text{F814W}<23.5$
sample that we want to use for our tests is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}
At $M_I\sim-22.0$, the
$m_\text{F814W}<23.5$ sample is beginning to lose galaxies in the
$0.9<z<1.0$ redshift bin due to the flux limit. However, the $0.85<z<1$
redshift bin was found to be only moderately overdense, so we choose
to disregard this region for the rest of the analysis, and instead
restrict to $z<0.85$, which is advantageous because it allows us to
choose a somewhat fainter intrinsic luminosity limit for the
analysis.  We relax our luminosity cut %to $-20.8$
so that the sample is volume-limited \emph{not} until $z=1$ but until $z=0.85$. 
We impose the cut at $M_I=-20.8$, which gives 95.3 per cent completeness in
the $0.8 < z \le 0.85$ bin (see Fig.~\ref{fig:MIhist}). The resulting
sample, which has 11~169 galaxies, will be called sample \s$1$ in the remainder
of this work.

However, previous studies
\citep[e.g.,][]{2003A&A...401...73W,2005ApJ...622..116G,2006ApJ...647..853W,Faber2007}
have shown that galaxy intrinsic luminosities evolve with
redshift. Thus, we should also let the luminosity cut that we apply to
volume-limit the sample evolve with redshift. 
Unfortunately, the majority of the published work on evolution of the
luminosity function uses $B$ and $V$ band data, and it is not apparent
that the results should be the same in a redder passband like $I$.  
We use the results from \cite{Faber2007} for the evolution of
$B$-band magnitudes from the DEEP2 and COMBO-17 surveys, which is $
\Delta M_B^* \sim -1.23$ mag per unit redshift (with the sign
indicating that galaxies were intrinsically brighter in the past), for a combined sample
of blue and red galaxies. 
Typically, estimates of evolution in the much redder $K$ band are less
than the estimates of evolution in $B$ and $V$ bands. (\rachel{Please
  include a citation for this statement.}) 
Assuming that the evolution is a smooth function of the wavelength,
the evolution in $I$-band should be in between $B$ and $K$ band. 
Therefore, by considering no evolution (a lower limit) as in our \s$1$, and a second
sample \s$2$ constructed using the $B$-band evolution (as an upper
bound on the $I$-band evolution), we can assume that these two samples
bracket reality, which should be somewhere in between.

Thus, \s$2$ is constructed by letting the luminosity cut evolve,
starting from $M_I = -20.8$ (same as in \s$1$) for the $0.8<z\le0.85$
bin.  The cut values for the other bins are defined by allowing $1.23$
magnitudes of evolution to fainter magnitudes as a function of
redshift (evaluated using the bin centers).  Because of the sign of
redshift evolution, \s$2$ includes  more galaxies.

One might wonder why we cannot use the luminosity function in F814W
based on the COSMOS observations to directly determine the luminosity
function for our sample, thus simplifying this exercise.  However,
this turns out to be highly non-trivial for two reasons.  First, the
F814W observations are relatively shallow compared to the deep
ground-based observations used in many other works for determination
of luminosity evolution.  As a result, it is difficult to get a handle
on the faint end of the luminosity function, and the unknown faint-end
slope turns out to be degenerate with the evolution of the typical
luminosity.  Second, the photometric redshift error is a complicating
factor that requires sophisticated techniques to remove.  A derivation
of the $I$-band luminosity evolution is therefore beyond the scope of
this work.

%In our analysis, we imposed the cut on luminosity for all redshift bins while many studies have shown that the characteristic luminosity of galaxies evolve with redshift. 
% The work that is at least close to what we expect is that of
% \cite{Faber2007}, who have studied the evolution of Schechter parameters in the B-band of DEEP2 and COMBO-17 surveys. We consider their reported value of $ \Delta M_B^* \sim -1.23$ per unit redshift, for both blue and red galaxies combined together.
%Although it is the $I$-band that is of relevance to us, we expect the evolution in B-band to be an upper bound on the evolution in I-band.

%Alternatively, we can impose cuts on stellar mass of the galaxies instead of imposing on the luminosity. 
%If we restrict to galaxies with masses $M$ such that $\log(M/M_\odot) \ge 10.15$, then we have at least {\bf 95.0\% } completeness.
Finally, we can circumvent the problem of redshift evolution of the
luminosity by imposing cuts on stellar mass instead. In Fig.~\ref{fig:smf}, we show the stellar mass function (SMF) of our sample for various F814W flux limits.
\cite{Tomczak_SMF} report the SMFs for the ZFOURGE survey, which
includes COSMOS. They considered for this work a single stellar
population (\rachel{What do you mean by a single stellar population?})
following a Chabrier IMF \citep{ChabrierIMF}. 
We plot their SMF for \emph{all} galaxies in Fig.~\ref{fig:smf} for
reference. Their SMF is higher than ours since they reach a $K_s$-band
$5\sigma$ depth of 24.9. 
As for $M_I$, we compare the stellar mass function in the $m_\text{F814W}\le23.5$ sample with that in the $m_\text{F814W}\le25.2$ sample. 
The sample with $\log(M/M_\odot) > 10.15$ is $\sim 95$ per cent
complete in the redshift bin $0.75 \le z< 0.85$ and has 10~341
galaxies in total across all redshifts.
Thus, we construct a third volume-limited sample \s$3$ by imposing the stellar mass cut mentioned above. The number of galaxies in redshift slices
are tabulated in Table~\ref{table:GalaxyCounts} for all three ways
discussed in this section for obtaining a volume-limited sample. The
stellar-mass limited sample is the smallest, most likely because when
converting from flux to stellar mass, the stellar mass-to-light ratios
vary strongly with galaxy type, so red galaxies with high $M_*/L$
simply have too low a flux compared to the blue galaxies at the same
$M_*$, and are not observed.

\begin{table} 
\centering
\begin{tabular}{|r|c|c|c|c|}
 \hline
 Redshift & Environment & \s$1$ & \s$2$ & \s$3$ \\
 \hline
 0.3-0.4 & Overdense & 1726 & 2505 & 1260 \\
 0.4-0.475 & Neutral & 988 & 1317 & 710 \\
 0.475-0.55 & Neutral & 1410 & 1788 & 902 \\
 0.55-0.65 & Underdense & 1797 & 2193 & 1183 \\
 0.65-0.75 & Overdense & 4059 & 4476 & 2593 \\
 0.75-0.8 & Underdense & 1159 & 1196 & 675 \\
 0.8-0.85 & Overdense & 2428 & 2428 & 1630 \\
 \hline
\end{tabular}
\caption{List of different redshift bins, their environmental
  classification and the number of galaxies per redshift bin for
  volume-limited samples constructed in three different ways: using a
  hard luminosity cut (\s$1$), using a redshift-dependent luminosity cut (\s$2$) and using stellar-mass cuts (\s$3$).}
\label{table:GalaxyCounts}
\end{table}

\begin{figure*}
 \centering
 \includegraphics[width=2.4\columnwidth]{SMF}
 \caption{Stellar mass distribution for various flux-limited samples
   are shown in three redshift ranges as separate panels. The redshift
   bins have been chosen to facilitate the comparison with  
          a study of the SMF in \protect\cite{Tomczak_SMF}. At high mass, the
          distributions are the same for various flux limits,
          indicating that the samples are complete in that mass
          range. The curves begin to deviate at low masses due to
          incompleteness coming from the flux limit.  The point at
          which the deviation exceeds our threshold determines where the mass cutoff should be to volume-limit the sample.}
 \label{fig:smf}
\end{figure*}

There is one subtlety in our method used for estimating
completeness. We have used the full $m_\text{F814W}\le23.5$ sample
for identifying
overdensities and for the completeness calculations that motivated our
definitions of volume-limited samples.  However, everywhere else in the paper, we consider only those galaxies for which there are postage stamp
images used to create weak lensing simulations, in part because this is the
sample for which fits to \sersic\ profiles were carried out, which is a
requirement for our morphology analysis.  
12 per cent of the galaxies that pass our cuts do not have an
associated postage stamp image.
Postage stamps may not exist because, given the size of the galaxy,
the size of the postage stamp we want to draw around it (including
some blank space) intersects the edge of the CCD.
If all galaxies were the same size, this would be a purely random effect, but in fact bigger galaxies are more likely to get excluded by this cut. 
It is commonly the case that  galaxies that are nearby and
intrinsically very bright do not have postage stamps associated with
them, an effect that is dominant at lower redshifts (and is part of
our reason for excluding $z<0.3$). 
Our completeness calculation is done at high redshifts, and thus we
believe that our conclusions are not affected by this bias.

%Not surprisingly, we see the overdense and underdense regions obtained by the above method have local average number density higher than the global average number density after volume limiting. %Refer to Figure~\ref{fig:comoving_densities}.
%And the environmentally neutral regions fall below or at least lie close to the 10\% margin.
%Refitting the model to our new sample is consistent with the previous assignment of overdensities.

The functional forms for the (flux-limited) redshift distribution that we used in Sec.~\ref{sub:overdensities} are not physically motivated. We fit them again to the  
redshift distribution of a volume-limited sample
(\s$1$). Fig.~\ref{fig:redshift_vollimited} shows that when we do
this, the values of $\delta_{g,\text{1D}}$ for the $z=0.40-0.55$ bin
increase and are within the $[-0.1, 0.1]$ range that we have defined
as neutral. This is the reason that in Sec.~\ref{sub:overdensities} we classified them as neutral as opposed to underdense.
We will see in Sec.~\ref{S:results} that they are more similar to overdense regions as opposed to underdense regions.
The other redshift slices seem to exhibit a consistent behavior in
Fig.~\ref{fig:redshift_vollimited} as in Fig.~\ref{fig:redshift_fluxlimited}.

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_vollimited}
  \caption{
Upper panel: Redshift distribution of volume-limited ($M_I <  -20.8$)
\s$1$ sample with photometric redshift bins that are 0.05 wide. Two
analytical functions with best fit parameters are plotted over it. 
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) =
           N/N_{\text{mod}}$ with each functional form as the model
           for each redshift bin. \rachel{I am quite concerned that
             something has gone wrong with this fit (in the sense that
           there may be a bug).  Why is the redshift distribution
           shown in blue forced to zero at $z=0.3$?  It should go to
           zero at $z=0$, not $0.3$.  It is clearly not normalized to the same total number as the redshift histogram, either.  In terms of formatting, please
           implement the same changes that I suggested for the
           flux-limited sample redshift distribution plot.}}
  \label{fig:redshift_vollimited}
\end{figure}

% \begin{figure}
%  \centering
%  \includegraphics[width=\columnwidth]{../plots_20140219/comoving_densities(2b).png}
%  \label{fig:comoving_densities}
%  \caption{In computing comoving densities, we have chosen the following values for the $\Lambda$CDM parameters: $\Omega_m = 0.3, \Omega_\Lambda = 0.7 \text{ and } \Omega_k = 0$ with the Hubble parameter $h = 0.72$.}
% \end{figure}

% \begin{tabular}{c|c|c|c|c|}
%  \hline
%  Redshift & \# of galaxies & Environment \\
%  \hline
%  0.30-0.35 & 726 & 1092 & Overdense \\
%  0.35-0.40 & 1000 1413 & & Overdense \\
%  0.40-0.45 & .. &  \\
%  0.45 - 0.50 & & \\
%  0.50-0.55 & & \\
%  0.55-0.60 & 727 & 905 & Underdense \\
%  0.60-0.65 & 1070 & 1288 & Underdense \\
%  0.65-0.70 & 2089 & 2321 & Overdense \\
%  0.70-0.75 & 1970 & 2155 & Overdense \\
%  0.75-0.80 & 1159 & 1196 & Underdense \\
%  0.80-0.85 & 2428 & 2428 & Overdense \\
%  \hline
% \end{tabular}

%In the following section, we will compare and analyze the distribution of properties of the galaxies residing in the overdense regions.

%\arun{Talk more about environments - Nature and nurture?}
%\rachel{What do you have in mind here?}

%\section{Implications/Comparisons}
%\section{Results}
\subsection{Describing galaxy morphology and shape}
\label{sub:axisratio}

% \rachel{As we discussed, you should get rid of the early parts of this
%   regarding second moments, and replace it with a brief description of
%   what shape estimates we use from GalSim (re-Gaussianization).}
% 
% {\bf Should explain in brief Claire's work.} \\

We choose simple and well-motivated ways to parametrize galaxy shapes
and morphology based on existing methods in the literature.  These
methods have the advantage of being stable and well-defined in nearly
all cases.  However, for highly irregular galaxies the meaning of the
structural parameters that we derive is not entirely clear.  In all
cases, our methods account for the effect of the {\em HST} PSF.
% Given the surface brightness (intensity) $I(\vec{\theta})$ of a galaxy image at every angular position $\vec{\theta}$,
% we can define the tensor of second brightness moments as
% \begin{equation}
%  Q_{ij} = \frac{\int\rmd^2\theta q_I (\theta_i - \overline{\theta_i})(\theta_j - \overline{\theta_j})}{\int\rmd^2\theta q_I},
% \end{equation}
% where $q_I$ is a suitable chosen kernel and $\overline{\vec{\theta}}$ is the average angular coordinate, weighted by the same kernel function and $i,j\in\lbrace 1,2 \rbrace$ are the components of $\vec{\theta}$.
% It is common to find in literature two kinds of \emph{complex }ellipticities that quantify the shape of the galaxies - $\chi$ and $\epsilon$ defined as
% \begin{align}
%  \chi &= \frac{Q_{11}-Q_{12}+2iQ_{12}}{Q_{11}+Q_{22}} \\ \epsilon &= \frac{Q_{11}-Q_{22}+2iQ_{12}}{Q_{11}+Q_{22}+2(Q_{11}Q_{22}-Q_{12})^{1/2}}
% \end{align}
% If the image has elliptical isophotes with axis ratio $q$, then 
% \begin{equation}
%  |\chi| = \frac{1-q^2}{1+q^2}, \;\;\; |\epsilon| = \frac{1-q}{1+q}. 
% \end{equation}
% 
% 
% Conversely, one could compute $\chi$ or $\epsilon$ for a galaxy, and obtain an effective, azimuthally averaged {\bf(?)} axis-ratio $q$. 
% \rachel{No, that isn't possible because of the PSF.  One would have to
%   use a PSF correction scheme.  Recommend removing much of the above discussion.}
% \arun{Should I remove these even if we plan to use the ellipticities from the second moments?}

One method to estimate the galaxy ellipticities and other
morphological parameters is to fit parametric models convolved with
the PSF to the observed galaxy light profile.  We use the fits from
\cite{great3}, which used the methods and
software from \cite{Claire_Fits} to fit the images to the following profiles:
\begin{enumerate}
 \item A \sersic profile given by the expression 
       \begin{equation} 
    I_S(x,y) = I_{1/2}\exp{\left[ -k(R(x,y)/R_{\text{eff}})^{1/n} -1 \right]},
       \end{equation}
where \begin{align*} R^2(x,y) = & ((x-x_0)\cos\Phi+(y-y_0)\sin\Phi)^2
  \\ & + ((y-y_0)\cos\Phi-(x-x_0)\sin\Phi)^2/q^2, \end{align*}
$R_{\text{eff}}$ is the half-light radius of the profile defined along
hte major axis, $I_0$ is the surface brightness at $R=R_{\text{eff}}$, $(x_0,y_0)$ is the centroid of the image,
$\Phi$ is the position angle, $n$ is the \sersic index, $k$ is a
$n$-dependent normalization factor required to ensure that half the
light is enclosed within the half-light radius, and $q$ is the axis
ratio of the elliptical isophotes.
Thus, the \sersic profile has 7 free parameters.
       \item A sum of two \sersic component fits: a de Vaucouleurs
         bulge $(n=4)$ plus an exponential disc profile $(n=1)$.  In this
         case, there are 10 free parameters, because the \sersic\
         indices are fixed, and the two components are constrained to
         have the same centroid.
\end{enumerate}
More details about the fitting algorithm can be found
in \cite{Claire_Fits}.

To quantify galaxy morphology and shape, we will use several
quantities from the above fits.  First, from the single \sersic
profile fits, we use the \sersic index and the axis ratio.  The axis
ratio can also be used to derive a distortion,
\begin{equation}\label{eq:distortion}
e = \frac{1-q^2}{1+q^2}
\end{equation}
or an ellipticity,
\begin{equation}\label{eq:ellipticity}
\varepsilon = \frac{1-q}{1+q}.
\end{equation}

As an alternative morphological indicator (instead of \sersic index)
we use a bulge-to-total ratio derived from the double \sersic
profile fits.  This ratio is defined in terms of bulge and disk fluxes as
\begin{equation}
\frac{B}{T} = \frac{f_\text{bulge}}{f_\text{bulge}+f_\text{disk}}.
\end{equation}
(\rachel{Note, the equation that was previously in here was wrong.
  It's not defined in terms of surface brightnesses, but rather in
  terms of flux.  Were you using numbers from my files for $B/T$?  If
  so, then it was calculated properly.  But if your analysis uses the
  equation that was in here before, then you will have to redo it.
  Sorry for not noticing this earlier.})

\rachel{I think it would be useful to take the entire sample that we
  use for science, and show the overall distributions of axis ratio,
  distortion (one curve for Claire's and one curve for
  re-Gaussianization), \sersic index, and \btt.  This would be a
  nearly full-page four panel figure, but I think it's worthwhile to
  illustrate the nature of the sample.  For example, it will clearly
  show why we cannot use the distributions of \sersic index and \btt,
  because of the hard cutoffs.}

We also consider an alternative method for estimating the galaxy
ellipticity.  This method is based on using the observed weighted
moments of the galaxy and PSF, and correcting those of the galaxy for
those of the PSF.  This PSF correction scheme is the
re-Gaussianization method described in section~2.4 of \cite{HS03}
(hereafter HS03) as implemented in the {\sc GalSim} software package
(with implementation details described in \citealt{2014arXiv1407.7676R}).
This method models the true PSF $g({\bf x})$ as a Gaussian $G({\bf x})$ and the residual $\epsilon({\bf x}) = g({\bf x}) - G({\bf x})$ is assumed to be small. Thus, the Gaussian-convolved
intrinsic image, $f$, can be modeled as $I' = G\otimes f = I - \epsilon \otimes f$, where $I$ is the observed image. The crucial idea here is that, when $\epsilon$ is small, we get a reasonably accurate
estimate of $I'$ even if we use an approximate form for $f$. The form
assumed for $f$ is that of a Gaussian with covariance $M_f^{(0)} = M_I
- M_g$, where $M_I$ and $M_g$ are the elliptical Gaussian-weighted adaptive
covariances of the measured object and PSF respectively, described in
section~2.1 of HS03 and \cite{BJ02}. We refer to the
re-Gaussianization estimates of the PSF-corrected distortion as
``moments-based shape estimates''.  The value in including them in
this analysis is that they have quite different radial weighting from
the \sersic profile fits, with the outer regions being quite
downweighted when calculating adaptive moments.  Thus, if ellipticity
gradients are important, we could get different results using these
two shape estimators.

\section{Results}
\label{S:results}

Having identified the overdense and underdense regions in a
volume-limited sample (Secs.~\ref{sub:overdensities}
and~\ref{sub:volumelimiting}), we will now see whether the morphological
parameters of the galaxies described in Sec.~\ref{sub:axisratio}
depend noticably on the environment of the redshift slice in which
they reside. Note that for true 3D overdensities there is already
substantial evidence in the literature that we should see variation of
properties with the environment.  Our test is necessary to see
whether such morphology-density correlations are evident in the kind
of 1D redshift slices that would be used for constructing weak lensing
simulations, or whether our use of an area the size of COSMOS will
wash out these trends (which would be good news for weak lensing
simulations based on that dataset).

As described in Sec.~\ref{sub:volumelimiting}, we have three different ways of volume-limiting our sample:
\begin{enumerate}
 \item no redshift evolution of luminosity cut $(\s1)$,
 \item using $B$-band luminosity evolution applied to the $I$-band
   luminosities $(\s2)$, and
 \item impose stellar mass cuts instead of luminosity $(\s3)$.
\end{enumerate}

We will present our results in all three cases to check for their
robustness to how the sample is selected.

\subsection{Axis ratios}

We can test the influence of environment on the galaxy shapes by
comparing the distributions of the axis ratios for the overdense and
underdense redshift slices, or by encapsulating that distribution as a
single number, the RMS (root mean squared) ellipticity. By
volume-limiting the sample, we have avoided issues
wherein the flux limit leads to artificial changes in the sample as a
function of redshift.  We will also carry out tests to differentiate
between environmental effects versus evolution of the population with
redshift (at fixed mass).

%Then, we proceed to consider the root mean square ellipticity, a statistic that can characterize the shape of the population/sample of galaxies. 
\subsubsection{Comparing distributions}\label{subsubsec:compdist}
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{axis_ratio_all}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_Bbandevolution.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_masscut.eps}
 \caption{The distributions of axis ratios of galaxies in \emph{all}
   overdense (OD) and \emph{all} underdense (UD) regions in the case
   of the luminosity-selected sample \s1 (left), luminosity-selected
   sample with $B$-band evolution taken into account \s2 (center), and
   thestellar-mass-selected sample \s3 (right). The upper panels show
   the histograms, 
 and the bottom panels show the cumulative distribution functions
 (CDF). The $p$-values computed using these CDFs are shown in Table. \ref{table:pvalues_all}.}
 \label{fig:axisratio_all}
\end{figure*}

\begin{figure*}
 \centering
 \includegraphics[width=2.3\columnwidth]{axis_ratio_odod}
 %a) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot3-0dot4_0dot65-0dot75.eps} \
 %b) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot3-0dot4_0dot80-0dot85.eps} \\
 %c) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot65-0dot75_0dot80-0dot85.eps} \
 %d) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot55-0dot65_0dot75-0dot8.eps} \
 \caption{Galaxy axis ratio
   distributions in two overdense redshift bins, $z=0.65-0.75$ and
   $z=0.80-0.85$, to check for consistency in the case that the
   environment is the same even if the redshift differs. The
   $p$-values from the KS and AD tests are given in
   Table.~\ref{table:pvalues_all}. \rachel{Figure does not appear to
     be centered.}}
 \label{fig:axisratio_similar}
\end{figure*}

\begin{figure*}
 \centering
 \includegraphics[width=2.3\columnwidth]{axis_ratio_odud}
 %a) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot55-0dot65_0dot65-0dot75.eps} \
 %#b) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot3-0dot4_0dot55-0dot65.eps} \\
 %c) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot3-0dot4_0dot75-0dot8.eps} \
 %d) \includegraphics[width=0.9\columnwidth]{axisratio(0)_0dot65-0dot75_0dot75-0dot80.eps} \
 \caption{Galaxy axis ratio distributions in a single underdense redshift
   slice, $z=0.55-0.65$, and a single overdense redshift slice,
   $z=0.65-0.75$. The $p$-values from the
   KS and AD tests are given in Table.~\ref{table:pvalues_all}.   \rachel{Figure does not appear to
     be centered.}}
 \label{fig:axisratio_contrasting}
\end{figure*}

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | c | c | c | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs.  & \scinot{1.1}{-4} & \scinot{2.6}{-5} & \scinot{1.9}{-6} \\
  All underdense     & \scinot{1}{-5} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.613 & 0.431 & 0.231 \\
  $[0.80,0.85]$ (OD) & 0.494 & 0.237 & 0.130 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{5.8}{-4} & \scinot{1.5}{-5} & \scinot{3.5}{-6} \\
  $[0.55,0.65]$ (UD) & \scinot{9.8}{-4} & $<\scinot{1}{-5}$ & $<\scinot{1}{-5}$ \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) tests obtained by comparing the
   distributions of axis ratios for three cases: \emph{all} overdense
   (OD) vs. \emph{all} underdense (UD), two overdense bins that are
   not very separated in redshift, and a pair of adjacent overdense
   and underdense bins. \s$1$, \s$2$, \s$3$ refer to the three
   different types of volume-limited samples.
          The Anderson-Darling $p$-values are correct only up to 5
          decimal places. 
\rachel{What does this last sentence mean?  Why?  I think you need
  clarification of this point in this table and the next one.}
% \rachel{What are the 0 entries in this and other tables?  Just how close to zero are they?  Since 0 is not really achieve, I recommend defining a value (such as $10^{-10}$ below which you will just say $<10^{-10}$ and give the value if it's above that, even if it seems really small to you, like $10^{-8}$.  Also, say $1.1\times 10^{-4}$ rather than $1.1e-4$.}
          }
 \label{table:pvalues_all}
\end{table}

\begin{table}
 \centering
 \begin{tabular}[\columnwidth]{ | l | l | l | l | }
  \hline
  %\parbox[t]{0.15\columnwidth}{Redshift bins} & \parbox[t]{0.15\columnwidth}{Tests} & \parbox[t]{0.15\columnwidth}{Luminosity-selected} & \parbox[t]{0.15\columnwidth}{Luminosity-selected with B-band evolution} & \parbox[t]{0.15\columnwidth}{Stellar-mass-selected} \\
  Redshift bins & \s$1$ & \s$2$ & \s$3$ \\
  \hline
  All overdense vs. & \scinot{5.6}{-4} & \scinot{1.0}{-4} & \scinot{3.3}{-6} \\
  All underdense    & \scinot{3}{-5} & \scinot{1}{-5} & $<\scinot{1}{-5}$ \\ \hline 
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & 0.9563 & 0.7476 & 0.5359 \\
  $[0.80,0.85]$ (OD) & 0.5162 & 0.3352 & 0.2290 \\ \hline
  $[0.65,0.75]$ (OD) vs.$\!\!\!$ & \scinot{6.0}{-3} & \scinot{2.5}{-4} & \scinot{2.4}{4} \\
  $[0.55,0.65]$ (UD) & \scinot{1.2}{-2} & \scinot{2.5}{-4} & \scinot{5}{-5} \\ \hline
 \end{tabular}
 \caption{$p$-values from the Kolmogorov-Smirnov (top) and
   Anderson-Darling (bottom) obtained by comparing the second
   moments-based ellipticities for three the same three cases as in
   Table~\ref{table:pvalues_all}.  \arun{Should I write only two
     significant digits for higher $p$-values?} \rachel{Yes.  The same
     comment applies to the previous table, too.}
           The Anderson-Darling $p$-values are correct only up to 5 decimal places. }
 \label{table:pvalues_momentbased_all}
\end{table}

%The distribution of the axis ratios that are obtained by the above manner are compared between redshift bins. 
We begin by comparing the entire axis ratio distributions $p(q)$
between pairs of redshift slices.  Unless otherwise mentioned, the
axis ratios refer to the values obtained using the method of
\cite{Claire_Fits} to fit single \sersic profiles to each galaxy image.
To compare the distributions and
make statistical statements about their consistency, we
use two statistical tests, the Kolmogorov-Smirnov (KS) test and
Anderson-Darling (AD) test, the latter of which is carried out using
the \texttt{adk} package in \texttt{R}.

We first compare the distribution of galaxy axis ratios in \emph{all}
overdense bins against that for \emph{all} underdense bins in
Fig.~\ref{fig:axisratio_all}, with different panels showing the
comparison for \s1, \s2, and \s3. 
The cumulative distinction functions are also shown, since they form
the basis for our statistical statements about consistency using the
KS and AD tests.  The results of these tests are shown in the first
two rows in Table~\ref{table:pvalues_all}.  For all three ways of
volume-limiting the sample, the $p$-values from both the KS and AD
tests are well below 0.05 (a maximum of $1.1\times 10^{-4}$, but often
smaller than that).  We can therefore reject the null hypothesis that
the overdense and underdense regions have the same underlying axis
ratio distributions at high significance.

One might imagine that the disagreement between the distributions is
at least partly due to the fact that the overdense and underdense
sample have different redshift distributions and there could be some
evolution of ellipticity distributions with redshift. 
To show that this redshift evolution effect is subdominant to
environmental effects, we will compare distributions between pairs of
two overdense (or pairs of underdense) redshift slices, where we
expect to find similarity even if the redshifts are different if the
environmental effects dominate.  We will also compare between
overdense and underdense regions that are selected to be nearby in
redshift, so that any redshift evolution effects should be minimal. 
Figures~\ref{fig:axisratio_similar}
shows that the axis ratio
distributions are indeed consistent when the environments are similar
but the redshifts are different.  Likewise
Fig.~\ref{fig:axisratio_contrasting} shows that for adjacent redshift
slices with different environments, the axis ratio distributions are
inconsistent.  The results of statistical tests for the distributions
in these figures are given in Table~\ref{table:pvalues_all}, and
support our statement that the morphology-density
correlation is the dominant effect when comparing overdense and
underdense redshift slices, with redshift evolution of the
population being negligible. 
Comparing other pairs of redshift
bins leads to similar conclusions. (\rachel{This argument would be even stronger if
  Fig.~\ref{fig:axisratio_similar} showed two redshift slices that were
  further separated in redshift, for example $0.3-0.4$ against
  $0.75-0.8$.  However the former might not have enough galaxies?})

Finally, we can check whether these findings are particular to the
axis ratios from the \sersic fits, or whether we reproduce this
finding when we use the shapes from the centrally-weighted
moments-based re-Gaussianization method, which estimates a distortion
(Eq.~\ref{eq:distortion}) for each galaxy. 
After neglecting a small fraction ($<0.01$ per cent) of galaxies for
which the method does not converge, 
we carry out the same statistical tests from
Table~\ref{table:pvalues_all}, but using the moments-based shape
estimates.  The results of the KS and AD tests are tabulated in Table~\ref{table:pvalues_momentbased_all}.
We see that all of our findings with the \sersic fit-based shapes
carry over to shapes from a centrally-weighted moments-based shape
estimate.

\subsubsection{RMS ellipticites}
\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \
% \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{RMS distortions as a function of redshift. The horizontal errorbars indicate
          the width of the redshift bin, while the vertical ones
          are $1\sigma$ errorbars obtained by
          bootstrapping. The colors and shapes of the points indicate their
          environmental classification, as shown in the legend.  \rachel{The title and figure clearly show two
          samples.  The caption needs to explain which are solid
          points and which are open. Also, the horizontal axis label should
          be either $z$ or ``Photometric redshift ($z$)'', not ``Photo
          $z$'' (which is not a phrase we have used throughout the
          paper).  The change in horizontal axis label is needed for
          the other RMS ellipticity plots and anywhere else that it
          shows up in the paper.}}
 \label{fig:rms_ellip}
\end{figure}

\begin{figure}
 %\centering
  \includegraphics[width=1.0\columnwidth]{rms_ellip1_Bbandevolution.eps} \
%\includegraphics[width=1.0\columnwidth]{rms_ellip2_Bbandevolution.eps} \\
 \caption{RMS distortions as a function of redshift. The solid
   points correspond to the sample where the luminosity cut evolves by
   $-1.23$ magnitudes per unit redshift (\s$2$) and the open
          points correspond to the sample obtained from stellar mass
          cuts (\s$3$). \rachel{If I compare this with
            Fig.~\ref{fig:rms_ellip} I would have concluded that the
            samples were the opposite of what is stated here.  The solid points appear the
            same in both plots, so I thought that was the stellar
            mass-selected sample.  Please check and confirm which way
            is correct, and include the information in both captions.}}
 \label{fig:rms_ellip_Bband_mass}
\end{figure}

\begin{figure}
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_noevolution.eps} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_Bbandevolution_masscut.eps} \\
 \caption{RMS distortions as a function of redshift, with points
   defined in a similar way as in Fig.~\ref{fig:rms_ellip}.  In this
   case, the distortions are the moments-based shape estimates, rather
   than from fits to \sersic profiles. }
 \label{fig:rms_ellip_momentbased}         
\end{figure}

\rachel{Overall, I find it quite confusing that the underdense regions
  are red and the overdense regions are blue on all of the plots in
  this section and the section on \sersic $n$ and $B/T$.
  The underdense regions are dominated by blue galaxies and the
  overdense are dominated by red galaxies, so having the plot colors
  reversed creates some cognitive dissonance.  Can you reverse them to
  more naturally match up point color with galaxy type?}

%The shape of a population of galaxies can be characterized by a single number called `RMS ellipticity'. If $q=b/a$ is the axis-ratio, then the ellipticity maybe defined as $\frac{1-q}{1+q}$
%or as $\frac{1-q^2}{1+q^2}$. The root mean squared (RMS) of ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}.
We can also carry out tests on a single statistic of the galaxy shape
distribution in each redshift slice, like the RMS ellipticity.  While
tests of a single quantity may seem less powerful than tests that use
the entire shape distributions, the advantage is that instead of
picking out pairs of redshift slices for our tests, we can easily compute our
statistic of interest for every single redshift slice, and look for
trends with both redshift and environment.

For the luminosity-selected sample without any evolution (\s1), the
RMS distortions (Eq.~\ref{eq:distortion}) of galaxies in each redshift bin are shown
in Fig.~\ref{fig:rms_ellip}.  (\rachel{The figure seems to show two
  samples, not just \s1.  Update the text to reflect this?}) For this and all similar figures, the
colors of the points were selected to easily differentiate between
galaxies in overdense, neutral, and underdense redshift slices.
As shown, the underdense regions
have  higher values for RMS ellipticities when compared to the
overdense regions. The difference between the underdense and overdense
regions for $z>0.5$ is significantly larger than any redshift
evolution across the $z>0.5$ range.  Our conclusions are very similar
if we use the RMS ellipticity from Eq.~\eqref{eq:ellipticity} instead
of the distortion. (\rachel{Our statements at the low redshift end
  regarding the amount of redshift evolution will depend on whether
  we're using the solid or open points, and the text and caption do
  not say which samples they correspond to, so I was not able to
  address this in the text.  Please do that once it is your turn to edit.})

The sign of the dependence on the local environment is reasonable when
compared with previous work on the morphology-density
relation. Overdense regions typically contain many old, elliptical
galaxies which are close to round (large axis ratio and low RMS ellipticity). 
In contrast, the underdense regions typically contain a larger
population of younger, disk galaxies, which have lower axis ratios and
higher RMS ellipticity. \rachel{Important note: this argument was
  given completely backwards.  I fixed it, but please watch out for
  this in future since it's a key point behind this work.}

From Figs.~\ref{fig:redshift_fluxlimited} and~\ref{fig:redshift_vollimited},
the $0.4\le z < 0.55$ redshift range shows signs of being marginally
underdense, but has somewhat low RMS ellipticities that agree with the rest of the overdense regions.

When the $B$-band luminosity evolution is taken into account in
selecting the sample, a systematic increase in the ellipticity at
lower redshifts can be observed. We plot these along with stellar-mass
selected samples, where a similar trend is found, in
Fig.~\ref{fig:rms_ellip_Bband_mass}.

Finally, we show an analogous plot of RMS ellipticities for all three
volume-limited sample using the moments-based shape estimates in
Fig.~\ref{fig:rms_ellip_momentbased}. The conclusions are quite
similar to those using the shapes from the \sersic profile fits, with
the underdense regions standing out in having larger RMS distortions 
than the other redshift slices, an effect that is substantially larger
than any average redshift evolution of the RMS distortions.

However, it is generally the case that the statistical significance of
trends in this section using a single statistic of the shape
distribution (the RMS) is less than the significance of the trends
seen using the entire axis ratio distributions in Sec.~\ref{subsubsec:compdist}.

\subsection{Other morphological parameters}

For the other morphological parameters that we described in
Sec.~\ref{sub:axisratio}, the \sersic index and bulge-to-total ratio,
we do not compare the distributions directly.  Doing so is relatively
tricky because both distributions have hard cutoffs that are enforced
in the fitting process (\sersic $n$ in the range $[0.25, 6]$ and
$0.05<B/T<0.95$).  The KS and AD statistical tests are both sensitive
to exactly what happens at these hard boundaries in the distributions.
So, instead of using the full distributions, we will study the
dependence of these quantities on environment by computing the median
values in different redshift slices.  Median values are preferred
over the full distributions or even the sample means since the medians
are more robust to what happens at the edges of the distributions.

Fig.~\ref{fig:median_sersicn} shows the median value of the \sersic
index in each redshift slice, with and without taking into account of
luminosity evolution when volume-limiting the samples (\s1 and
\s2). Both panels also show the results with stellar mass-selected
samples (\s3) for reference.  
When using the stellar mass-selected sample, we observe that the
overdense regions tend to have higher \sersic 
index than the underdense ones, with the redshift evolution being mild
and the results for the underdense regions particularly standing
out. This trend is consistent with our previous explanation for trends
in RMS ellipticities; the underdense regions have more spiral galaxies
and therefore a lower median value of \sersic $n$. 
However, this trend is less evident for the luminosity-selected
samples, where there seems to be some evolution with redshift that
dominates over the environmental effects.

We also note that in Fig.~\ref{fig:median_sersicn}, the median \sersic
indices of the stellar mass-selected samples (\s$3$) are
systematically greater than those of the luminosity selected samples
(\s$1$, \s$2$).  We saw that \s3 is the smallest sample, and thus is
likely limited to higher mass galaxies overall.  It is therefore not
surprising that it has a higher median \sersic $n$.  Finally, even for
the stellar mass-selected sample there is some sign of redshift
evolution.  The sign of this evolution is as expected, with lower
\sersic $n$ and $B/T$ for higher redshift samples, which should have a
higher fraction of disk and irregular galaxies and fewer galaxies with
bulge-like morphology.

Finally, Fig.~\ref{fig:median_dvcbtt} shows the variation of the median
bulge-to-total ratio with redshift. The results are quite consistent
with those of Fig.~\ref{fig:median_sersicn}. 
Thus, our results in this section suggest that the 
environment can significantly affect the median morphological
parameters of galaxies selected in thin redshift
slices, assuming that the galaxies represent a stellar mass-selected
sample.  The trend is less evident when using luminosity to select the
galaxies.

%We chose median over mean since \sersic n greater than 6 are truncated to 6. This affects the mean statistic more than the median.
\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{median_sersicn}
 \includegraphics[width=1.0\columnwidth]{median_sersicn2}
 \caption{Median values of the \sersic indices for volume-limited
   samples \s$1$ and \s$3$ are plotted (filled centers and thin
   errorbars) in top and bottom panels, respectively, for each
   redshift bin. 
          Median values for the \s$2$ sample are plotted in both the
          panels (open centers and thick errorbars) in both the
          panels.
\rachel{Now I'm really confused.  The text implies
     that the top and bottom are \s1 and \s2 as open points, with \s3 in both
     panels as solid points.  The caption says something different
     regarding which sample is in both panels and which is open
     vs. closed.  Please clarify and make everything consistent!
     Also, you only need to have a legend on one panel, since the
     legends on both panels are the same.  Same comment applies to
     next figure.}}
 \label{fig:median_sersicn}
\end{figure}

\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{median_dvcbtt1}
 \includegraphics[width=1.0\columnwidth]{median_dvcbtt2}
  \caption{\rachel{Same confusion here as in previous figure caption;
      please clarify!} Median values of the bulge-to-total ratios for volume-limited samples \s$1$ and \s$3$ are plotted (filled centers and thin errorbars) in left and right panels respectively for each redshift bin.
          Median values for the \s$2$ sample are plotted in both the panels (open centers and thick errorbars) in both the panels.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones are $1\sigma$ errorbars obtained by bootstrapping.  \rachel{These look so similar that I am questioning the need to have both.  Perhaps remove one and mention that they are nearly identical in the text.}}
 \label{fig:median_dvcbtt}
\end{figure}

\section{Conclusions and implications}
\label{S:summary}

\rachel{Here is one thing that is still missing from the results
  section.  We had discussed a number of times lumping the galaxies
  together into slices with larger $\Delta z$ and showing that the
  trends in the shape distributions disappear or are reduced.  You had
  shown me plots for this at some point, but they don't seem to have
  gone into the paper.  Please include them, because it would be
  useful to be able to give some guidance on how wide the redshift
  slices should be to avoid this issue.}

In this study, we have shown that the shape distributions of galaxies
(to a statistically significant degree) and morphological parameters
like \sersic $n$ and bulge-to-total ratios (more marginally) depend on
the local environments when dividing up the COSMOS sample into
redshift slices along the line of sight.  The redshift slices used had
a width of $\Delta z=0.05$.  Our findings are robust to the choice of
shape estimator from \sersic profile fits vs. using centrally weighted
moments-based shear estimates.

These findings are relevant to attempts to use {\em HST}-based galaxy
samples to calibrate shear estimates in weak lensing surveys.  In
general, the approach would be to define galaxy samples using all
galaxies in redshift slices, and determine a redshift-dependent shear
calibration.  Our findings highlight the danger in such an approach:
while we would like our simulations to include true evolution in
galaxy properties with redshift, this approach also includes spurious
variations in galaxy properties due to the large-scale structure
within the COSMOS field.  Since the fidelity of weak lensing shear
estimates depends sensitively on the intrinsic shape distribution and
galaxy morphological parameters, the conclusions for the
redshift-dependent shear calibration would be incorrect.  These errors
would be reduced as the redshift slices that are used become wider, so
that the impact of local overdensities becomes washed out.

It is important to keep in mind the nature of COSMOS with respect to
other possible {\em HST} training samples.  COSMOS represents the
largest contiguous field surveyed by the {\em HST}, with the sizes of
other {\em HST} fields such as GOODS and AEGIS lagging by at least an
order of magnitude.  Hence, if cosmic variance due to structures along
the line-of-sight in COSMOS are problematic for its use as a training
sample for weak lensing simulations, studies that use even smaller
area training samples are even more prone to errors, with the UDF
serving as an extreme case.  If the size of the {\em HST} survey is
small enough, there is no reason {\em a priori} to suppose that the
galaxy population is typical even when using all galaxies along the
line of sight, without any division into redshift bins.  Of course,
future surveys are unlikely to pick just a single survey to serve as
the basis for their image simulation training sample, but rather
to combine as many as possible.  Combining multiple surveys will
reduce the cosmic variance and therefore the significance of the
effects discussed in this work.  However, with COSMOS being so
dominant in its size, the combination with other much smaller fields
is unlikely to fully ameliorate this effect.  Thus, it will be
important to carefully choose the size of redshift slices used to
derive properties of the galaxy population so as to be minimally
affected by this issue.

It is natural to ask the question of how relevant are variations in
the galaxy population of the size found here for shear calibration
bias estimates.  It is likely that the answer to this question varies
quite significantly with the type of shear estimation method used.
Some will be sensitive to the variations in morphology, others to the
variation in the intrinsic ellipticity distribution, and many will be
sensitive to both at some level.  We consider each in turn.

The intrinsic ellipticity distribution plays a role in nearly all
shear estimators.  In some, the role is explicit: for example, LensFit
\citep{2007MNRAS.382..315M,2008MNRAS.390..149K,2013MNRAS.429.2858M}
and the methods presented by \cite{2014MNRAS.438.1880B} require
accurate intrinsic ellipticity distributions as inputs, and
uncertainty in the distribution was one of the important sources of
systematic uncertainty in the CFHTLenS weak lensing results
\citep{2013MNRAS.432.2433H,2013MNRAS.429.2858M}.  In others, the role
is implicit.  For example, the re-Gaussianization method and several
other moment-based methods require calculation of a shear responsivity
\citep{BJ02,HS03} that describes how the galaxy population overall
responds to a shear, based on its intrinsic ellipticity distribution.
The responsivity can be calculated based on the observed shape
distribution, assuming that the uncertainties in the shears are known
enough that their contribution to that distribution can be removed.
If a simulated sample in some redshift slice has a different intrinsic
ellipticity distribution and therefore responsivity, it could lead to
incorrect conclusions about shear calibration.  The responsivity
scales roughly like $1-e_\text{RMS}^2$, which means that deviations in
RMS ellipticity at the level of 0.01 would become fractional shear
errors of
\begin{equation}
\frac{\Delta\gamma}{\gamma} \approx \frac{2 e_\text{RMS} \Delta
  e_\text{RMS} }{1-e_\text{RMS}^2} \approx 0.01.
\end{equation}
In the context of upcoming lensing surveys that seek to constrain
shears to the per cent level or better, a systematic error of this
magnitude in shear calibration is quite serious.

Regarding possible biases in the morphological mixtures of galaxies
due to overdensities or underdensities in the training sample, there
are results in the literature for several methods that show how shear
biases vary with morphology.  For example, for the maximum likelihood
fitting code {\sc im3shape}, figure 2 in \cite{2012MNRAS.427.2711K}
shows multiplicative biases for two-component \sersic profile galaxies
as a function of their bulge-to-total ratios (denoted there as
$F_b/(F_b+F_d)$, which we will equate with our $B/T$).  The shear
calibration bias scales roughly like $0.04 - 0.05(B/T)$ as $B/T$ goes
from $0$ to $1$.  Our results suggest that typical (median) $B/T$
values may be influenced by cosmic variance in the COSMOS field
leading to fluctuations of order $0.05$.  The resulting variation in
the calibration bias would therefore be $\sim 2.5\times 10^{-3}$, or
$0.25$ per cent shear calibration uncertainty.  For existing datasets
this is not very problematic, but for surveys like LSST, Euclid, and
WFIRST-AFTA, this would be a dominant part of the systematic error
budget.  To give another example, for re-Gaussianization, figure 9 of
\cite{2012MNRAS.420.1518M} shows that as \sersic $n$ goes from $1$ to
$6$, the shear calibration bias varies by $2$ per cent.  In this case,
since we have shown that the median value of \sersic $n$ can vary by
$\sim 0.4$ due to morphology-density correlations, this suggests that
the shear calibration for re-Gaussianization could be misestimated by
$\sim 2\times 10^{-3}$, or 0.2 per cent.  This too is acceptable in
existing datasets, but not those that will be used for shear
estimation in the next decade.

A final consideration is the question of how applicable these results
using volume-limited samples are to simulations of upcoming weak
lensing surveys, which will exclusively use flux-limited samples.  For
our analysis, the volume-limiting sample was necessary to avoid
complications due to varying galaxy populations in each redshift
slice, allowing us to isolate purely environmental effects.  In
principle, if the morphology-density correlations that we have
identified turn out to not exist for intrinsically fainter galaxy
populations, then at low redshift (where a flux limited sample will
include galaxies that are intrinsically much fainter than at high
redshift), the effects will be less serious for upcoming lensing
surveys.  However, we do not have any particular reason to believe
that these effects will vanish for fainter galaxies.  Moreover, at
higher redshift where only intrinsically bright galaxies can be seen,
the effect should be present at a level similar to what we have found
here.  Since higher redshift galaxies tend to dominate cosmological
shear estimates (due to their higher shears), our findings will be
important to take into account.

In conclusion, our results have serious implications for the plans to
create realistic image simulations that will be used to derive
redshift-dependent shear calibrations for upcoming weak lensing
surveys.  If care is not taken to mitigate this effect, then the
cosmic variance in the training sample may bias the conclusions
regarding shear calibration for redshift slices that represent
significant overdensities or underdensities compared to the typical
galaxy population.  This is particularly a problem when using the
smaller {\em HST} surveys, where a single galaxy cluster or a void
could completely dominate the galaxy population in a given redshift
slice.  To mitigate this problem, it will be imperative to (a) collect
training data from widely separated patches on the sky, and (b) take
care to use redshift slices that are brought enough that these effects
are reduced, so as to wash out the effect of any signal overdensity or
underdensity on the simulated galaxy population.  By employing these
mitigation schemes, there is every reason to believe that the effect
we have identified can be reduced to a small component of the
systematic error budget of major upcoming lensing surveys.

\section*{Acknowledgments}

AK and RM acknowledge the support of NASA ROSES 12-EUCLID12-0004, and
program HST-AR-12857.01-A, provided by NASA through a grant from the
Space Telescope Science Institute, which is operated by the
Association of Universities for Research in Astronomy, Incorporated,
under NASA contract NAS5-26555. RM acknowledges the support of an Alfred P. Sloan Research Fellowship.  We thank Alexie Leauthaud for 
many useful discussions.

\rachel{Need to fix arXiv references so that they show up properly
  with e-print number.}

%\begin{thebibliography}
\bibliography{ref}
%\end{thebibliography}
\end{document}